### 2022-03-31 manuscript.Rmd
figure compiled

---
title: "Reproducible and programmatic generation of neuroimaging visualisations"
author: "Sidhant Chopra, Löic Labache, Winnie Orchard, [...]"
date: "09/03/2022"
output:
  pdf_document: default
  word_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(rgl)
rgl::setupKnitr()
rgl::rgl.open(); rgl::rgl.close() # first window in doc. may not render properly
```

### Introduction

Visualising neuroimaging data is one of the primary ways in which we
evaluate data quality, interpret results and communicate findings. These
visualisations are commonly produced using graphic interface-based (GUI)
tools where individual images are opened and display settings are
manually changed until the desired output is reached. While the choice
to use GUI-based software's has been driven by a perception of
convenience, flexibility and accessibility, there are now many
code-based software packages which are well-documented and don't often
require high-level knowledge of programming. These tools are flexible
and allow for generation of reproducible, high-quality and publication
ready brain visualisations in only a few lines of code (Figure 1),
especially within the R, Python and Matlab environments.

This article will cover three major advantages of using code-generated
visualisations over GUI based tools: replicability, flexibility and
integration. First, by generating figures using code, you increase the
replicability of your figures for yourself, your collaborators and your
readers. Second, code-based software provide more precise controls over
visualisation settings, while also benefiting from advantages of
programming, such as being able to iteratively generate multiple
figures. Finally, being able to integrate figure generation into your
analysis scripts reduces chances of errors, with more advanced tools now
allowing for seamless weaving together of prose and code. Here, we
review the advantages of learning and using programmatic neuroimaging
visualisations, focusing on benefits to replicability, flexibility and
interactivity. We conclude by introducing examples of R-packages which
allow for visualisation of region-of-interest (ROI), voxel, vertex and
edge level data, followed by a brief discussion of limitations and
functionality gaps in code-based brain visualisation tools.

### Replicability

In recent years, there have been multiple large-scale efforts
empirically demonstrating the lack of reproducibility of findings from
neuroimaging data (cite). One common solution proposed for achieving
robust and reliable discoveries has been to encourage scientific output
which can be transparently evaluated and potentially replicated. In
practise, what this usually means is openly sharing detailed methods,
materials, code and data. While there is a trend towards increasing
transparency and code sharing in neuroimaging research (cite), the
sharing of code used to generate figures which include brain renderings
and spatial maps has been relatively neglected. This gap in
reproducibility is partly driven by the fact that brain figures are
often created using a manual process of tinkering with sliders, buttons
and overlays on a GUI, concluding with a screenshot and sometimes
beatification in image processing software like Illustrator. This
process inherently makes neuroimaging visualisations difficult, if not
impossible to replicate, even by the authors who created them.

Given that brain figures regularly form the centrepiece of
interpretation within a paper, conference presentations and news
reports, making sure they can be reliably regenerated is crucial for
knowledge generation. By writing and sharing code used to generate brain
visualisations, a direct and traceable link is established between the
underlying data and scientific figure. While the code that produces a
replicable figure doesn't reflect the validity of the scientific finding
or the accuracy of the content figure, it instils transparency and
robustness while demonstrating a desire to further scientific knowledge.
Some would even consider publishing figures which cannot be replicated
as closer to advertising, rather than science (cite blog).

### Flexibility

Being able to exactly replicate your figures via code has serious
advantages beyond good open science practises. In particular, the
ability to reprogram inputs (such as statistical maps) and settings
(such as colour schemes and visual orientations) can streamline your
entire scientific workflow. Being able to code the inputs and settings
allows for the easy production of multiple figures, such as those
resulting from multiple analyses which require similar visualisations.
Likewise, an arduous request from a reviewer to alter the data
processing or analysis becomes less of a burden when the figure can be
re-generated with a few lines of code, as opposed to re-pasting and
re-illustrating them. Having a code-base with reprogrammable inputs can
mean that the generation of visualisations requires less time, energy
and effort than manual GUI-based generation. Keep in mind that the gains
of writing code for your figures are cumulative, and in addition to
improving your programming, you start to build a codebase for figure
generation that you can continue to reuse and share through your
scientific career.

While brain visualisations are often thought of as the end results of
analyses, they also form a vital part of quality control of imaging
data. While tools to automatically detect artefacts, denoise the data
and generate derivatives are becoming more robust, we are not yet at the
stage where visualising the data during processing is no longer need.
Nonetheless, when working with large datasets such as Human Connectome
Project (cite) or UK BioBank (cite), it is infeasible to use traditional
GUI based tool to visually check the data. Knowing how to
programmatically generate brain visualisations can allow you to iterate
your code over each image of a large datasets making quality controls of
each data processing or analysis step accessible and achievable.

### Integrative and Interactive reporting

Often programming languages such as R, Python and Matlab are used for
the analysis in neuroimaging studies, but the brain visualisation
resulting from these analyses is outsourced to a septate GIU-based
visualisation tools such as FSLeyes, Freeview or ITK-snap. Switching
from analysis to a GUI-based visualisation process be a cumbersome
deviation from the scientific workflow and can make debugging errors
more difficult, as you have to regularly switch program to visually
examine the results of any modifications to analyses. Using the brain
visualisation tools that already exist within your chosen programming
environment can provide instant visual feedback on the impact of
modifications to processing or analyses. While some GUI-based tools do
offer command-line access to generate visualisations, they can lack
flexibility to easily generate publication ready figures and can lose
the benefits provided by your preferred programming environment.

Increasingly popular tools such as R-Markdown, Quarto and Jupyter
notebooks allow for the mixing of prose and code in a single script,
resulting in fully reproducible and publication ready papers. By using
code-based tools available within your preferred environment, brain
visualisations can be directly integrated and embedded within a paper or
report. Journals that publish neuroimaging studies are increasingly
moving towards allowing reproducible manuscripts, including reproducible
figures (cite e-life,f1000,plos), with with some even allowing on-demand
re-running of code using cloud services (cite).

Neuroimaging data are often spatially 3D and can have multiple time
points, adding a 4th dimensional (e.g. fMRI data). Thus, communicating
findings or evaluating quality using static 2D slices is challenging.
While well-curated 3D renderings can help with spatial localisation
(cite madan), in the end, static images can only provide an incomplete
representation of the data. A added advantage of some of code-based
tools is that you can generate 'rich' media like interactive figures or
animations, which can improve scientific communication of findings. Many
code-based software tools now allow for the generation of animated brain
visualisations (cite blog) or interactive widgets which allow users to
zoom, rotate and scroll through slices. Linking to or even embedding
these videos or interactive figures in papers can greatly enhance the
communication of findings.

### Examples of neuroimaging visualisation packages available in R

##### *Voxel-level visualisations.*

The `oro.nifti` package allows for the reading, writing, manipulation
and visualisation of voxel-level imaging data of either *nifti*,
*analyze* or *afni* formats. In particular, the `image`, `slice` and
`orthographic` functions allow users to display the desired number of
slices in the desired orientations, while also providing precise control
over overlaid images, colour scales, legend placement and other
aesthetic settings (Fig 1A). Smoothing and re-sampling of images is
sometimes required for visualisation purposes, which can be achieved via
the *`ANTsR`* /`extrantsr` or `fslr` packages.

[view_xifti_volume(xii, interactive = TRUE) in cifti tools]

##### *Vertex-level visualisations.*

The recently developed `fsbrain` package allows for the visualisation of
vertex-level and atlas-level data obtained using Freesurfer. It contains
many flexible functions for visualising individual-level and group-level
measures such as cortical thickness, volume or surface area (Fig 1B).
The package also comes with extensive guides and documentations to
assist users (cite).

CIFTI or 'grey-ordinate' data is becoming a popular format for
structural and functional imaging data, as it combines vertex-level data
of the cortex with voxel-level data fo the cerebellum and sub-cortex. In
R, CIFTI data can now be read, visualised and manipulated using the
well-documented `ciftiTools` package (Fig 1C; cite).

##### *ROI-level visualisations.*

Statistical parameters can be mapped onto discreet cortical and
sub-cortical regions using the `ggseg` and `ggseg3d` package. These
packages flexibly generates aesthetic renderings of cortical,
sub-cortical and white matter ROIs in 2D and 3D (Fig 1D). Usually,
ROI-level interpretations are dependent on a standardised atlas or
parcellation schemes the brain, accordingly this package and its add-ons
(`ggExtra`) provides a large array of commonly used atlases and some
functionality for users to contribute other atlases. Additionally, being
part of the 'Grammar of Graphics' frame work allows for integration with
packages such as `gganimate`, enabling users to animate and visualise
dynamic changes in spatial and temporal statistics across ROIs. The
`ciftiTools` package also allows for surface and voxel-level
visualisation of ROIs for the cortex and sub-cortex respectively (Fig
1E). Flexibly allowing CIFTI format atlases and ROIs to be displayed.

##### Edge-level visualisations.

Analyses which divide the brain into discrete regions and examine
pair-wise dependences between regional phenotypes are becoming
increasingly common. Often the results of these dependency analyses are
statistics relating to 'edges' or links between any two regions. The
package `brainconn` allows users to plot brain regions as nodes of a
network graph, and dependences between regions as edges of that graph,
in both 2D and interactive 3D (Figure 1D). Similar to `ggseg`, this
package comes within multiple commonly used atlases and contains
functionality for users to enter their own atlas schemes.

### Limitations and functionality gaps in R packages for brain visualisation

-   While most of these tools do not require a strond knowledge of
programming, there can still be a steeper learning curve when
compared to GUI-based tools.

-   For quick and interactive viewing of single images, sometimes GUI
tools can be faster (although, fsleyes can be called from python and
R environments with ease)

-   Visualising data-types such as streamlines from DWI-based
tractography are still not well represented in code-based tools.

-   But fsbrain has a function;
<https://github.com/dfsp-spirit/fsbrain/blob/cb1b9d667a9449b214b20f6420038af2541e48b4/R/vis_dti_tracks.R>)

-   For tools which require non-standard or uncommon atlases or ROIs.

-   Although, with `ciftitools`, this becomes easier, as long as
atlas follows the CIFTI convention.

-   3D renderings of non-standard ROIs like unique subcortical schemes
or brain stem are still difficult to do.

-   Have to use complex tools like pyVista

### Figure 1

```{r Figure1 }
library(patchwork)
library(ggplot2)

fig1a <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1a.png"),
                                                       width=ggplot2::unit(1,"npc"),
                                                       height=ggplot2::unit(1,"npc")))

fig1b <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1b.png"),
                                                       width=ggplot2::unit(1,"npc"),
                                                       height=ggplot2::unit(1,"npc")))

fig1c <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1c.png"),
                                                       width=ggplot2::unit(1,"npc"),
                                                       height=ggplot2::unit(1,"npc")))

fig1d <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1d.png"),
                                                       width=ggplot2::unit(1,"npc"),
                                                       height=ggplot2::unit(1,"npc")))

fig1e <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1e.png"),
                                                       width=ggplot2::unit(1,"npc"),
                                                       height=ggplot2::unit(1,"npc")))

fig1f <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1f.png"),
                                                       width=ggplot2::unit(1,"npc"),
                                                       height=ggplot2::unit(1,"npc")))

fig1g1 <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1g1.png"),
                                                        width=ggplot2::unit(1,"npc"),
                                                        height=ggplot2::unit(1,"npc")))
fig1g2 <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1g2.png"),
                                                        width=ggplot2::unit(1,"npc"),
                                                        height=ggplot2::unit(1,"npc")))

fig1h <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1h.png"),
                                                       width=ggplot2::unit(1,"npc"),
                                                       height=ggplot2::unit(1,"npc")))

fig1i <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG("data/fig1i.png"),
                                                       width=ggplot2::unit(1,"npc"),
                                                       height=ggplot2::unit(1,"npc")))

patchwork <- fig1a/fig1b|(fig1c|fig1f)/fig1d|(fig1e|fig1g2)/fig1g1|fig1h/fig1i

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualisations in R')

```

```{r Fig1a, eval=FALSE}
#Figure 1A
#voxel 1

# ADD COMMENTS TO EXPLAIN EACH LINE

library(neurobase)
library(ggplotify)

template <- readnii('data/MNI152_T1_1mm_brain.nii.gz') #load the nifti file into R

effect <- readnii("data/MNI152_effect_size.nii.gz")

#threshold the effect at the desired cut-off
effect[effect<0.3] <- NA

#Plot the effect on the brain template
breaks = seq(.3,.6, by=0.005)

png("data/fig1a.png") 
ortho2(x = template, 
       y = effect, 
       crosshairs = F, 
       bg = "white", 
       NA.x = T, 
       col.y= hotmetal(), 
       xyz = c(70,50,80),
       useRaster = T,
       ycolorbar = TRUE, 
       mfrow = c(1,3)) +
  colorbar(breaks,col = hotmetal(n=length(breaks)-1), 
           text.col = "black", labels = TRUE, maxleft = 0.95) + 
  title(main = "Voxel Statistic Map")
dev.off()

```

```{r Fig1b, eval=FALSE }
#Figure 1B
#voxel 2

library(neurobase)
library(scales)

dti <- readnii('data/sub-001__t1_warped.nii.gz') #load the nifti file into R

atlas <-  readnii('data/sub-001_schaefer300n7_aseg_to_dwispace_gm_rois.nii.gz')

png('data/fig1b.png')
overlay(x=robust_window(dti), y=atlas, 
        plot.type = "single", 
        z=c(seq(30,110,10)), 
        col.y =  alpha(rainbow(300), 0.3),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, 
        zlim.y =c(1,300), 
        mar = c(0.1, 0.1, 1.5, 0.1)) + #marget of bottom, left, top, right borders
  title("Voxel Atlas Overlay", line = 1.3)
dev.off()

```

```{r Fig1c , eval=FALSE}
#vertex 1
#Figure 1B
library(rgl)
library(ggplot2)
# Load the package and point to the Connectome Workbench
suppressPackageStartupMessages(library(ciftiTools))

ciftiTools.setOption("wb_path", "/Applications/workbench/")

# Read and visualize a CIFTI file

cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
surfL_fname <- ciftiTools.files()$surf["left"]

cii <- read_cifti(
  cifti_fname, brainstructures="left", 
  surfL_fname=surfL_fname)

view_xifti_surface(cii, 
                   hemisphere = "left", 
                   view = "both", 
                   idx = 1,
                   colors = viridis::viridis(n = 100),
                   zlim = c(1,1.5),  
                   legend_embed = T,
                   title = "Vertex Statistic Map", 
                   cex.title = 2)

rgl.snapshot("data/fig1c.png")
rgl.close()

```

```{r Fig1d , eval=FALSE}
#Vertex 2
#Figure 1X - Remove white space

library(fsbrain)

#download_optional_data()
subjects_dir = get_optional_data_filepath("subjects_dir")

subject_id = 'subject1'

colourmap = colorRampPalette(viridis::viridis(n = 100))

vis.subject.morph.native(subjects_dir = subjects_dir, 
                         subject_id = 'subject1', 
                         measure = 'thickness', 
                         hemi='both', 
                         views = 't4',
                         draw_colorbar = "horizontal",
                         cortex_only = T, surface = "white", 
                         makecmap_options = list('colFn'=colourmap))
rgl.snapshot("data/fig1d.png")
rgl.close()

```

```{r Fig1e , eval=FALSE}
#ROI1 
#Figure 1e 
suppressPackageStartupMessages(library(ciftiTools))
parc <- load_parc("Yeo_7")
view_xifti_surface(parc,  
                   legend_embed = T,
                   hemisphere = "right")

rgl.snapshot("data/fig1e.png")

rgl.close()

```

```{r Fig1f , eval=FALSE}
#ROI1 
#Figure 1X - combine with previous figure

set.seed(1993) #set a random seed (good practice for reproducibility)

parc <- load_parc("Schaefer_400")

ramdom_metric <- rnorm(400)

cii <- move_from_mwall(cii, NA)

parc_vec <- c(as.matrix(parc))

xii_metric <- c(NA, ramdom_metric)[parc_vec + 1]

xii1 <- select_xifti(cii, 1)

xii_metric <- newdata_xifti(xii1, xii_metric)

plot2 <- view_xifti_surface(xii_metric, 
                            colors = viridis::viridis(n = 400), 
                            borders = "black", 
                            hemisphere = "left", )

rgl.snapshot("data/fig1f.png")

rgl.close()

```

```{r Fig1g, eval=FALSE}
#ROI2

#devtools::install_github("LCBC-UiO/ggsegGlasser")
library(ggseg)
library(ggplot2)
library(ggsegGlasser)

set.seed(1993) #set a random seed (good practice for reproducibility)

base_atlas <- as.data.frame(na.omit(cbind(glasser$data$region, glasser$data$hemi)))

colnames(base_atlas) <- c("region", "hemi")

Effect <- rnorm(dim(base_atlas)[1]) #generate a random numbers for each roi in the atlas

base_atlas <- cbind(Effect, base_atlas)

cortex <- ggseg(atlas = glasser,
                .data = base_atlas,
                mapping=aes(fill=Effect), 
                position="stacked",
                colour="black",
                size=.2,
                show.legend = F,
                plot.background = "white") + scale_fill_viridis_b()

sub_base_atlas <- as.data.frame(na.omit(cbind(aseg$data$region, aseg$data$hemi)))

colnames(sub_base_atlas) <- c("region", "hemi")

Effect <- rnorm(dim(sub_base_atlas)[1])

sub_base_atlas <- cbind(Effect, sub_base_atlas)

subcortex <- ggseg(atlas = aseg,
                   .data = sub_base_atlas,
                   hemisphere = "right",
                   mapping=aes(fill=Effect), 
                   position="stacked",
                   colour="black",
                   size=.2,
                   show.legend = T, 
                   plot.background = "white") + 
  scale_fill_viridis_b()

#fig1g <- cowplot::plot_grid(cortex, subcortex, nrow = 1, rel_widths = c(1,.7))

ggsave(filename = 'data/fig1g1.png', plot = cortex, device = 'png', bg = 'white')
ggsave(filename = 'data/fig1g2.png', plot = subcortex, device = 'png', bg = 'white')

```

```{r Fig1h , eval = F}
#Edge (also add 3d version)
#Fig1X

#remotes::install_github("sidchop/brainconn")
library(brainconn)

conmat <- example_weighted_undirected
degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]
fig1h <- brainconn(atlas ="schaefer300_n7", 
                   conmat=conmat, 
                   node.size = degree/2,
                   view="ortho", 
                   edge.color.weighted = T,
                   background.alpha = 0.4, 
                   show.legend = F,
                   edge.color = scale_edge_colour_gradient2(low='yellow', mid='orange', high='red')) 
ggsave(filename = 'data/fig1h.png', plot = fig1h, device = 'png', bg = 'white')

brainconn3D(atlas ="schaefer300_n7", 
            conmat=conmat,
            show.legend = F)

```