### 2022-04-05 manuscript.Rmd
fig update

---
title: "Reproducible and programmatic generation of neuroimaging visualisations"
author: "Sidhant Chopra, Löic Labache, Winnie Orchard, [...]"
date: "09/03/2022"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

### Introduction

Visualising neuroimaging data is one of the primary ways in which we
evaluate data quality, interpret results and communicate findings. These
visualisations are commonly produced using graphic interface-based (GUI)
tools where individual images are opened and display settings are
manually changed until the desired output is reached. The choice to use
GUI-based software's has been driven by a perception of convenience,
flexibility and accessibility. However now there are also many
code-based software packages which are well-documented and don't often
require high-level knowledge of programming, making them more accessible
to the neuroimaging community. These tools are flexible and allow for
the generation of reproducible, high-quality and publication-ready brain
visualisations in only a few lines of code (Figure 1), especially within
the R, Python and Matlab environments.

This article will cover three major advantages of using code-generated
visualisations over GUI based tools: replicability, flexibility and
integration. First, by generating figures using code, you increase the
replicability of your figures for yourself, your collaborators and your
readers. Second, code-based software provides more precise controls over
visualisation settings, while also benefiting from advantages of
programming, such as being able to iteratively generate multiple
figures. Finally, being able to integrate figure generation into your
analysis scripts reduces chances of errors, with more advanced tools now
allowing for seamless weaving together of prose and code. Here, we
review the advantages of learning and using programmatic neuroimaging
visualisations, focusing on benefits to replicability, flexibility and
intergration. We conclude by introducing examples of R-packages which
allow for visualisation of statistics at region-of-interest (ROI),
voxel, vertex and edge level data, followed by a brief discussion of
limitations and functionality gaps in code-based brain visualisation
tools.

### Replicability

In recent years, there have been multiple large-scale efforts
empirically demonstrating the lack of reproducibility of findings from
neuroimaging data (cite poldrack). One common solution proposed for
achieving robust and reliable discoveries has been to encourage
scientific output which can be transparently evaluated and independently
replicated. In practise, what this usually means is openly sharing
detailed methods, materials, code and data. While there is a trend
towards increasing transparency and code sharing in neuroimaging
research (cite), the sharing of code used to generate figures which
include brain renderings and spatial maps has been relatively neglected.
This gap in reproducibility is partly driven by the fact that brain
figures are often created using a manual process of tinkering with
sliders, buttons and overlays on a GUI, concluding with a screenshot and
sometimes beatification in image processing software like Illustrator.
This process inherently makes neuroimaging visualisations difficult, if
not impossible to replicate, even by the authors who created them.

Given that brain figures regularly form the centrepiece of
interpretation within a paper, conference presentations and news
reports, making sure they can be reliably regenerated is crucial for
knowledge generation. By writing and sharing code used to generate brain
visualisations, a direct and traceable link is established between the
underlying data and scientific figure. While the code that produces a
replicable figure doesn't reflect the validity of the scientific finding
or the accuracy of the content figure, it instils transparency and
robustness while demonstrating a desire to further scientific knowledge.
Some would even consider publishing figures which cannot be replicated
as closer to advertising, rather than science (cite blog).

### Flexibility

Being able to exactly replicate your figures via code has serious
advantages beyond good open science practises. In particular, the
ability to reprogram inputs (such as statistical maps) and settings
(such as colour schemes and visual orientations) can streamline your
entire scientific workflow. Changing inputs and settings via code allows
for the easy production of multiple figures, such as those resulting
from multiple analyses which require similar visualisations. While some
GUI-based tools do offer command-line access to generate visualisations,
they can lack flexibility to easily generate publication ready figures
and can lose the benefits, such as iteration, provided by your preferred
programming environment. A simple for-loop or copying and pasting the
code with the input and/or setting-of-interest modified can be a
powerful method for exploring visualisation options or creating
multi-panel figures. Likewise, an arduous request from a reviewer to
alter the data processing or analysis becomes less of a burden when the
figure can be re-generated with a few lines of code, as opposed to
re-pasting and re-illustrating them. Having a code-base with
reprogrammable inputs can mean that the generation of visualisations
requires less time, energy and effort than manual GUI-based generation.
This also makes it easier to make figures for each subsequent project
that you work on. Keep in mind that the gains of writing code for your
figures are cumulative, and in addition to improving your programming,
you start to build a code-base for figure generation that you can
continue to reuse and share through your scientific career.

Precise controls via code over visualization settings, such as colour
schemes, legend placement and camera angles, can provide you with much
greater flexibility over visualisations. Nonetheless, part of the appeal
of GUI-based tools is that the pre-sets for such settings can provide a
useful starting point and reduce the decision burden on the user.
However, similar pre-sets are often available in the from of default
settings in most code-based packages, negating the need for the user to
manually enter each and every choice of setting required for making the
image. Most code-based tools also come with detialed documentation, or
even entire papers (cite ggseg and ciftitools) which provide examples of
figures which can be used as starting points or templates for new users.
As the popularity of code-sharing for figure increases, there will be an
cornucopia of templates that can be used as starting points for new
figures.

While brain visualisations are often thought of as the end results of
analyses, they also form a vital part of quality control of imaging
data. Tools to automatically detect artefacts, de-noise the data and
generate derivatives are becoming more robust, but we are not yet at the
stage where visualising the data during processing is no longer need.
Nonetheless, when working with large datasets such as Human Connectome
Project (cite) or UK BioBank (cite), it is unfeasible to use traditional
GUI-based tool to visually check the data. The time it takes to open a
single file and achieve the desired visualisation settings vastly
compounds when working with large datasets. Knowing how to
programmatically generate brain visualisations can allow you to iterate
your visualisation code over each image of a large datasets making
checking the quality of each data processing or analysis step accessible
and achievable. The visual outputs of each iteration can be complied
into accessible documents which can be easily scrolled, with more
advanced usage allowing for the creation of interactive HTML reports,
similar to those created by tools like mriQC (cite). Increasing capacity
to conduct visual quality control on larger datasets can increase the
identification of processing errors and result in more reliable and
valid findings from your data.

### Integrative and Interactive reporting

Often programming languages such as R, Python and Matlab are used for
the analysis in neuroimaging studies, but the brain visualisation
resulting from these analyses is outsourced to separate GIU-based
visualisation tools such as FSLeyes, Freeview or ITK-snap. Switching
from your analysis environment to a GUI-based visualisation process be a
cumbersome deviation from the scientific workflow. This can make
debugging errors more difficult, as you have to regularly switch program
to visually examine the results of any modifications to analyses. Using
the brain visualisation tools that already exist within your chosen
programming environment can provide instant visual feedback on the
impact of modifications to processing or analyses.

Increasingly popular tools such as R-Markdown, Quarto and Jupyter
Notebook allow for the mixing of prose and code in a single script,
resulting in fully reproducible and publication ready papers. By using
code-based tools available within your preferred environment, brain
visualisations can be directly integrated and embedded within a paper or
report. Some journals that publish neuroimaging studies are moving
towards allowing reproducible manuscripts, including reproducible
figures (cite e-life,f1000,plos), with some even allowing on-demand
re-running of code using cloud services (cite).

Neuroimaging data are often spatially 3D and can have multiple time
points, adding a 4th dimension (e.g. fMRI data). Thus, communicating
findings or evaluating quality using static 2D slices is challenging,
and may not be the best representation of the data, or the
interpretations. While well-curated 3D renderings can help with spatial
localisation (see madan), in the end, static images can only provide an
incomplete representation of the data, and forces researchers to choose
the best angle to show, which often involves compromising one result for
another. An added advantage of some of the code-based tools is that you
can generate 'rich' media like interactive widgets -- figures or
animations, which allow users to zoom, rotate and scroll through slices.
Interacting with a figure in this way can improve scientific
communication of findings. Linking to or even embedding these videos or
interactive figures in papers can greatly enhance the communication of
findings and make your paper more engaging for the reader. Such rich
brain visualisations lend themselves to being embedded or shared on
science communication mediums other than academic papers - such as
presentations, web-sites and social media - all of which can promote the
communication of your research with peers and a reach larger audiences.

### Examples of neuroimaging visualisation packages available in R

The following four sections provide brief examples packages and
functions available in R for visualising voxel, vertex, ROI and
edge-level data. A didactic introduction to these tools is provided in a
accompaning [collection of beginner friendly
R-Notebooks](ttps://github.com/sidchop/neuRo-vis-BrainHack-2021).

```{=tex}
\newpage
\blandscape
```
### Figure 1 [Fix image size ratios]

```{r Figure1 , echo=FALSE, out.height = '80%', out.width = '140%'}
library(patchwork)
library(ggplot2)

fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1g3','fig1h', 'fig1i')
figs <- list()

for (f in 1:length(fig_name)) { 
  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/',fig_name[f],'.png')),
                                                             width=ggplot2::unit(1,"npc"),
                                                             height=ggplot2::unit(1,"npc")))
}

patchwork <- figs[[1]]/(figs[[2]]|figs[[3]])|(figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[4]]|figs[[10]]/figs[[11]]

#& theme(text = element_text('Times New Roman'))

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualisations in R')

```

\
*[Figure 1 - Add Caption and remove panel F]*

\elandscape

##### Voxel-level visualisations.

The `oro.nifti` package allows for the reading, writing, manipulation
and visualisation of voxel-level imaging data of either nifti, analyze
or afni formats. In particular, the `image`, `slice` and `orthographic`
functions allow users to display the desired number of slices in the
desired orientations, while also providing precise control over overlaid
images, colour scales, legend placement and other aesthetic settings
(Fig 1A-B). Smoothing and re-sampling of images is sometimes required
for visualisation purposes, which can be achieved via the `ANTsR`
/`extrantsr` or `fslr` packages.

##### Vertex-level visualisations.

The recently developed `fsbrain` package allows for the visualisation of
vertex-level and atlas-level data which is derieved using FreeSurfer
formats. It contains many flexible functions for visualising
individual-level and group-level measures such as cortical thickness,
volume or surface area (Fig 1E). The package also comes with extensive
guides and documentations to assist users (cite).

CIFTI or 'grey-ordinate' data is becoming a popular format for
structural and functional imaging data, as it combines vertex-level data
of the cortex with voxel-level data for the cerebellum and sub-cortex.
In R, CIFTI data can now be read, visualised and manipulated using the
well-documented `ciftiTools` package (Fig 1C-D,F; cite).

##### ROI-level visualisations.

Statistical parameters can be mapped onto discreet cortical and
sub-cortical regions using the `ggseg` and `ggseg3d` package. These
packages flexibly generates aesthetic renderings of cortical,
sub-cortical and white matter ROIs in 2D and 3D (Fig 1D). Usually,
ROI-level interpretations are dependent on a standardised atlas or
parcellation schemes the brain, accordingly this package and its add-ons
(`ggExtra`) provides a large array of commonly used atlases and some
functionality for users to contribute other atlases. Additionally, being
part of the 'Grammar of Graphics' frame work allows for integration with
packages such as `gganimate`, enabling users to animate and visualise
dynamic changes in spatial and temporal statistics across ROIs.

The `ciftiTools` package also allows for surface and voxel-level
visualisation of ROIs for the cortex and sub-cortex respectively (Fig
1E). Flexibly allowing CIFTI format atlases and ROIs to be displayed.

##### Edge-level visualisations.

Analyses which divide the brain into discrete regions and examine
pair-wise dependences between regional phenotypes are becoming
increasingly common. Often the results of these dependency analyses are
statistics relating to 'edges' or links between any two regions. The
package `brainconn` allows users to plot brain regions as nodes of a
network graph, and dependences between regions as edges of that graph,
in both 2D and interactive 3D (Figure 1D). Similar to `ggseg`, this
package comes within multiple commonly used atlases and contains
functionality for users to enter their own atlas schemes.

### Limitations and functionality gaps in R packages for brain visualisation

-   While most of these tools do not require a strong knowledge of
    programming, there can still be a steeper learning curve when
    compared to GUI-based tools.

-   For quick and interactive viewing of single images, sometimes GUI
    tools can be faster (although, fsleyes can be called from python and
    R environments with ease)

-   Visualising data-types such as streamlines from DWI-based
    tractography are still not well represented in code-based tools.

-   Lack of user friendly tools which can render non-standard or
    uncommon atlases or ROIs. Although, with `ciftitools`, this becomes
    easier, as long as atlas follows the CIFTI convention.

-   3D renderings of non-standard ROIs like unique subcortical schemes
    or brain stem are still difficult to do. Can to use complex
    workarounds like ITKsnap + pyVista.

### Supplement (Code used to generate figure)

```{r Fig1a, eval=FALSE}
#Figure 1A
#voxel 1

# ADD COMMENTS TO EXPLAIN EACH LINE
library(neurobase)
library(ggplotify)

template <- readnii('data/MNI152_T1_1mm_brain.nii.gz') #load the nifti file into R

effect <- readnii("data/MNI152_effect_size.nii.gz")

#threshold the effect at the desired cut-off
effect[effect<0.3] <- NA

#Plot the effect on the brain template
breaks = seq(.3,.6, by=0.005)

png("data/fig1a.png") 
ortho2(x = template, 
       y = effect, 
       crosshairs = F, 
       bg = "white", 
       NA.x = T, 
       col.y= viridis::viridis(n = 500), 
       xyz = c(70,50,80),
       useRaster = T,
       ycolorbar = TRUE, 
       mfrow = c(1,3)) +
  colorbar(breaks,col =viridis::viridis(n=length(breaks)-1), 
           text.col = "black", labels = TRUE, maxleft = 0.95) 
dev.off()

```

```{r Fig1b, eval=FALSE }
#Figure 1B
#voxel 2

library(neurobase)
library(scales)

dti <- readnii('data/sub-001__t1_warped.nii.gz') #load the nifti file into R

atlas <-  readnii('data/sub-001_schaefer300n7_aseg_to_dwispace_gm_rois.nii.gz')

png('data/fig1b.png')
overlay(x=robust_window(dti), y=atlas, 
        plot.type = "single", 
        z=c(seq(30,110,10)), 
        col.y =  alpha(rainbow(300), 0.3),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, 
        zlim.y =c(1,300))

      #  mar = c(0.1, 0.1, 1.5, 0.1)) + #marget of bottom, left, top, right borders
dev.off()

```

```{r Fig1c , eval=FALSE}
#vertex 1
#Figure 1B
library(rgl)
library(ggplot2)
# Load the package and point to the Connectome Workbench
library(ciftiTools)

ciftiTools.setOption("wb_path", "/Applications/workbench/")

# Read and visualize a CIFTI file

cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
surfL_fname <- ciftiTools.files()$surf["left"]
surfR_fname <- ciftiTools.files()$surf["right"]

cii <- read_cifti(
  cifti_fname, brainstructures="all", 
  surfL_fname=surfL_fname,
  surfR_fname=surfR_fname)

view_xifti_surface(cii, 
                   hemisphere = "left", 
                   view = "both", 
                   idx = 1,
                   colors = viridis::viridis(n = 100),
                   zlim = c(1,1.5),  
                   legend_embed = T,
                   cex.title = 2)

rgl.snapshot("data/fig1c.png")
rgl.close()

```

```{r Fig1d , eval=FALSE}
#Vertex 2
#Figure 1X - Remove white space

library(fsbrain)

#download_optional_data()
#download_fsaverage(accept_freesurfer_license = TRUE)

subjects_dir <- get_optional_data_filepath("subjects_dir")

subject_id <- 'subject1'

colourmap <- colorRampPalette(viridis::viridis(n = 10000))
fsbrain.set.default.figsize(700, 700);

cm = vis.subject.morph.standard(subjects_dir, 
                                subject_id, 'sulc', 
                                fwhm='10', 
                                cortex_only = T, 
                                views=NULL, 
                                makecmap_options = list('colFn'=colourmap))
img = export(cm, colorbar_legend='Effect', output_img = "data/fig1d.png")

rgl.snapshot("data/fig1d.png")
rgl.close()

```

```{r Fig1e , eval=FALSE}
#ROI1 
#Figure 1e 
suppressPackageStartupMessages(library(ciftiTools))
parc <- load_parc("Yeo_7")
view_xifti_surface(parc,  
                   legend_embed = T,
                   hemisphere = "right")

rgl.snapshot("data/fig1e.png")

rgl.close()

```

```{r Fig1f , eval=FALSE}
#ROI1 
#Figure 1X - combine with previous figure

set.seed(1993) #set a random seed (good practice for reproducibility)

parc <- load_parc("Schaefer_400")

ramdom_metric <- rnorm(400)

cii <- move_from_mwall(cii, NA)

parc_vec <- c(as.matrix(parc))

xii_metric <- c(NA, ramdom_metric)[parc_vec + 1]

xii1 <- select_xifti(cii, 1)

xii_metric <- newdata_xifti(xii1, xii_metric)

plot2 <- view_xifti_surface(xii_metric, 
                            colors = viridis::viridis(n = 400), 
                            borders = "black", 
                            hemisphere = "both",
                            view = 'lateral')

rgl.snapshot("data/fig1f.png")

rgl.close()

```

```{r Fig1g, eval=FALSE}
#ROI2

#devtools::install_github("LCBC-UiO/ggsegGlasser")
library(ggseg)
library(ggplot2)
library(ggsegGlasser)

set.seed(1993) #set a random seed (good practice for reproducibility)

base_atlas <- as.data.frame(na.omit(cbind(glasser$data$region, glasser$data$hemi)))

colnames(base_atlas) <- c("region", "hemi")

Effect <- rnorm(dim(base_atlas)[1]) #generate a random numbers for each roi in the atlas

base_atlas <- cbind(Effect, base_atlas)

cortex <- ggseg(atlas = glasser,
                .data = base_atlas,
                mapping=aes(fill=Effect), 
                position="stacked",
                colour="black",
                size=.2,
                show.legend = F,
                plot.background = "white") + scale_fill_viridis_b()

sub_base_atlas <- as.data.frame(na.omit(cbind(aseg$data$region, aseg$data$hemi)))

colnames(sub_base_atlas) <- c("region", "hemi")

Effect <- rnorm(dim(sub_base_atlas)[1])

sub_base_atlas <- cbind(Effect, sub_base_atlas)

subcortex <- ggseg(atlas = aseg,
                   .data = sub_base_atlas,
                   mapping=aes(fill=Effect), 
                   position = "dispersed",
                   colour="black",
                   hemi =c('left', 'right'),
                   size=.2,
                   show.legend = T, 
                   plot.background = "white") + 
  scale_fill_viridis_b()

ggsave(filename = 'data/fig1g1.png', plot = cortex, device = 'png', bg = 'white')
ggsave(filename = 'data/fig1g2.png', plot = subcortex, device = 'png', bg = 'white')

```

```{r, eval=FALSE}

make_ggseg3d <- function(attribute, 
                         colour.pal = c("light yellow",
                                        "orange",
                                        "red",
                                        "dark red"), 
                         hide.colourbar=FALSE, 
                         output.png=FALSE, 
                         file.name="ggseg_3d.png") {
  # Inputs: 
  #' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left-Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Right-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Right-Amygdala, Right-Accumbens-area

  library(ggseg3d)
  library(tidyr)
  library(dplyr)
  #remove(aseg_3d)
  aseg_3d <- aseg_3d
  aseg_3d <- tidyr::unnest(aseg_3d, cols = c(ggseg_3d))

  attribute.ggseg3d <- c(rep(NA, 4), attribute[1:4], rep(NA, 3), attribute[5:7],
                         rep(NA, 5), attribute[8:14],  rep(NA, 6))

  data <- dplyr::mutate(aseg_3d, attribute =  attribute.ggseg3d)
  #remove NA regions
  aseg_3d[which(is.na(data$attribute)),] <- NA
  aseg_3d <- tidyr::drop_na(aseg_3d)

  data[which(is.na(data$attribute)),] <- NA
  data <- tidyr::drop_na(data)
  data$attribute[data$attribute==0]<-NA #make 0 valus NA so they are set as grey in ggseg3d

  scene=list(camera = list(eye = list(x = 0, y = 1, z = -2.25)),
             aspectratio = list(x=1.6,y=1.6,z=1.6))

  plot <- ggseg3d::ggseg3d(.data = data, 
                           atlas = aseg_3d, 
                           colour = "attribute", 
                           text = "attribute",
                           palette = colour.pal)
  plot <- remove_axes(plot)
  plot <- plotly::layout(plot, 
                         scene = scene, 
                         width = 600, height = 600)

  if(hide.colourbar==TRUE){
    plot <- plotly::hide_colorbar(plot)
  }

  if( output.png==TRUE){
    plotly::orca(plot, file =  file.name)
  }

  return(plot)

}

set.seed(1993)
volume <- sample(1:20, 14) #randomly generated data for 14 structures
make_ggseg3d(volume, 
             colour.pal = viridis::viridis(n=length(volume)), 
             output.png = T, file.name = "data/fig1g3.png")

```

```{r Fig1h , eval = F}
#Edge (also add 3d version)
#Fig1X

#remotes::install_github("sidchop/brainconn")
library(brainconn)

conmat <- example_weighted_undirected
degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]
fig1h <- brainconn(atlas ="schaefer300_n7", 
                   conmat=conmat, 
                   node.size = degree/2,
                   view="ortho", 
                   edge.color.weighted = T,
                   background.alpha = 0.4, 
                   show.legend = F,
                   edge.color = scale_edge_colour_gradient2(low='yellow', mid='orange', high='red')) 
ggsave(filename = 'data/fig1h.png', plot = fig1h, device = 'png', bg = 'white')

brainconn3D(atlas ="schaefer300_n7", 
            conmat=conmat,
            show.legend = F)

```